The Influence of Connectivity on Dung Beetle Communities

### 2025-07-25 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
doc formating

 # - doublespacing # manipulate line spacing 
# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
# - \usepackage{xcolor} # changes colors of text
# # http://latexcolor.com/ to choose colors
# # define the colors used above with rgb codes
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# # to center to top level (#) headings (my name) and change color of font
# # more info # https://www.ctan.org/pkg/sectsty
# - \usepackage{sectsty}
# - \sectionfont{\centering\color{darkmidnightblue}}
# # to center to sub-sub-headings (###) and change color of font
# - \usepackage{sectsty} 
# - \subsectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} 
# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
# - \usepackage{wrapfig} 
# - \newcommand{\blandscape}{\begin{landscape}}
# - \newcommand{\elandscape}{\end{landscape}}
# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
# - \usepackage{indentfirst}
# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
bibliography: ["eric-thesis.bib"]
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: true
    fig_caption: true
classoption: 12pt
# output: bookdown::word_document2
# output: word_document
# output: html_document

@@

Although richness and diversity were the same among treatments, there was notable variability between sampling blocks. Block 8 generally had the lowest species richness and biodiversity while 53n had the highest. While the experimental design attempted to control for the effects of patch size and edge, there could be large (and potential unknown) environmental gradients across the SRS landscape that could influence the observed patterns in diversity and abundance. For instance, at the time of my study Block 8 had the densest matrix of any of the blocks. This could have hindered the diffusion of bait scent, leading to lower capture rates in this block. Other habitat characteristics that might dsiffer among blocks could have been influential as well -- for example, soil quality and forest cover can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019]. The same is true for land-use history; much of the SES land would have previously been used for agriculture, and during development of experimental units cleared with heavy machinery might experience heavy soil compaction *brudvig*. Landscape features could also affect mammal movement, which might in turn limit dung availability - Block 8 was nearest to roadways used by all employees at SRS, and the eastern side of the site had large bodies of water upon which mammals are dependent [@WOS:000241362400032; @dechen_quinn_landscape_2013; @WOS:000651446900001]. Additionally, deer and feral hogs are the most dominant mammals at the site, two species that do well in patchy landscapes [@castillo-contreras_urban_2018; @fraser_connectivity_2019]. The locally variable patterns of abundance in the landscape of these generalist large mammals, coupled with ability to easily move through modified landscapes, might help explain the limited variation in richness or diversity within blocks (i.e., between patch types) but large block effects. Deer also commonly forage along the edges of habitat patches and forest ecotones. If they are spending more time in these locations, the higher abundance and diversity in connected and winged patches might in part be due to  'drift-fence' effects **citation of levey drift fence**.

>>>>

 # - doublespacing # manipulate line spacing 
# - \usepackage{parskip} # manipulate indents and spacing before/after paragra phs 
# - \usepackage{xcolor} # changes colors of text
# # http://latexcolor.com/ to choose colors
# # define the colors used above with rgb codes
# - \definecolor{darkcerulean}{rgb}{0.03, 0.27, 0.49}
# - \definecolor{darkmidnightblue}{rgb}{0.0, 0.2, 0.4}
# # to center to top level (#) headings (my name) and change color of font
# # more info # https://www.ctan.org/pkg/sectsty
# - \usepackage{sectsty}
# - \sectionfont{\centering\color{darkmidnightblue}}
# # to center to sub-sub-headings (###) and change color of font
# - \usepackage{sectsty} 
# - \subsectionfont{\centering\color{darkmidnightblue}}
# - \usepackage{sectsty} 
# - \subsubsectionfont{\raggedright\color{darkmidnightblue}}
# - \usepackage{wrapfig} 
# - \newcommand{\blandscape}{\begin{landscape}}
# - \newcommand{\elandscape}{\end{landscape}}
# - \usepackage[font=small,labelfont=bf]{caption} #Required for specifying captions to tables and figures
# - \usepackage{indentfirst}
# - \setlength\parindent{22pt} # uncomment if you want indented paragraphs
bibliography: ["eric-thesis.bib"]
#output:
#  bookdown::pdf_document2:
#    toc: false
#    number_sections: true
#    fig_caption: true
#classoption: 12pt
output: bookdown::word_document2
# output: word_document
# output: html_document

@@

Although richness and diversity were the same among treatments, there was notable variability between sampling blocks. Block 8 generally had the lowest species richness and biodiversity while 53n had the highest. While the experimental design attempted to control for the effects of patch size and edge, there could be large (and potential unknown) environmental gradients across the SRS landscape that could influence the observed patterns in diversity and abundance. For instance, at the time of my study Block 8 had the densest matrix of any of the blocks. This could have hindered the diffusion of bait scent, leading to lower capture rates in this block. Other habitat characteristics that might differ among blocks could have been influential as well -- for example, soil quality and forest cover can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019]. The same is true for land-use history; much of the SES land would have previously been used for agriculture, and during development of experimental units cleared with heavy machinery might experience heavy soil compaction *brudvig*. Landscape features could also affect mammal movement, which might in turn limit dung availability - Block 8 was nearest to roadways used by all employees at SRS, and the eastern side of the site had large bodies of water upon which mammals are dependent [@WOS:000241362400032; @dechen_quinn_landscape_2013; @WOS:000651446900001]. Additionally, deer and feral hogs are the most dominant mammals at the site, two species that do well in patchy landscapes [@castillo-contreras_urban_2018; @fraser_connectivity_2019]. The locally variable patterns of abundance in the landscape of these generalist large mammals, coupled with ability to easily move through modified landscapes, might help explain the limited variation in richness or diversity within blocks (i.e., between patch types) but large block effects. Deer also commonly forage along the edges of habitat patches and forest ecotones. If they are spending more time in these locations, the higher abundance and diversity in connected and winged patches might in part be due to  'drift-fence' effects **citation of levey drift fence**.
